Risks of Running Vulnerable Container Images for Data Analysis on HPC Clusters

### 2019-07-31 paper.tex
july 31

No doubt containers have gained a lot of popularity after Docker, however containers tightly
integrate with the host due to the sharing of the kernel \TG{not only the kernel: 
mounted file systems and devices too}, which raises security
concerns. Figure \ref{fig:container-overview} provides overview of the containers architecture.
They also share three main security threats with other virtualization
techniques~\cite{gantikow2016providing}.

@@

\textbf{Privilege Escalation} A malicious attacker can break out of the container
and gain control of the host system and other containers running on the same host.

\textbf{Denial-of-Service} One container may end up eating all the resources of the
host hence resulting in the starvation of the host and other containers.

\textbf{Information Leak} Private data of the host and other containers could be
leaked and can be used for further attacks.

\TG{Avoid bold face in papers}

However, two features of the Linux kernel, \textit{control groups} and \textit{namespaces},
are used by containers to mitigate these threats to a certain extent. Security
of containers is still most concerning issue.
Except web applications, containers are also popularly used for scientific computing
and data science in neuroscience. In fact, Singularity was developed for scientific applications
by keeping focus on HPC environments.
The goal of this paper is to report all vulnerabilities that are present in
container images, specifically used by neuroimaging field, and access their
related risks for data analysis. \TG{You should explain the 3 questions discussed during out meeting.}

@@

Docker images are organized as a series of layers stacked on top of each
other and identified by unique layer identifiers \TG{If a layer is reused
in different images, for instance the top layer of a base image, will it
have the same id across images?}. Each layer consists of a filesystem diff
with respect to the layer below. Layers are compressed \TG{Are you
referring to file compression similar to gzip?} into a single image and can
be pushed to DockerHub. There are two ways to create a Docker image. The
first is to use an existing Docker image as a base image and then build a
new image from that one, by making changes in the filesystem, which is
saved as a new layer of the Docker image. The second way is to build a new
Docker image from scratch. \TG{It would be relevant to mention the file
system used to store the images (is it squashfs?).}

>>>>

No doubt containers have gained a lot of popularity after Docker, however containers tightly
integrate with the host due to the sharing of the kernel, 
mounted file systems, and devices, which raises security
concerns. Figure \ref{fig:container-overview} provides overview of the containers architecture.
They also share three main security threats with other virtualization
techniques~\cite{gantikow2016providing}.

@@

\textit{Privilege Escalation} A malicious attacker can break out of the container
and gain control of the host system and other containers running on the same host.

\textit{Denial-of-Service} One container may end up eating all the resources of the
host hence resulting in the starvation of the host and other containers.

\textit{Information Leak} Private data of the host and other containers could be
leaked and can be used for further attacks.

However, two features of the Linux kernel, \textit{control groups} and \textit{namespaces},
are used by containers to mitigate these threats to a certain extent. Security
of containers is still most concerning issue. According to a Forrester survey~\cite{} done in January
2015, security is the
biggest concern while deploying containers.
Except web applications, containers are also popularly used for scientific computing
and data science in neuroscience. In fact, Singularity was developed for scientific applications
by keeping focus on HPC environments. This popularity containers lead us to the
primary research question of this work: \textit{what is the current security state of
container images that are deployed on HPC clusters}.

To answer this question, we take a real use case of containers into consideration.
The goal of this paper is to report all vulnerabilities that are present in
container images, specifically used by neuroimaging field, and access their
related risks for data analysis. Particularly, we aim to answer three primary
research questions:

\textit{RQ1} Are vulnerabilities present inside container images that are
deployed on HPC clusters?
\textit{RQ2} Can we get rid of these vulnerabilities by updating the
images?
\textit{RQ3} Can we get rid of these vulnerabilities by shrinking the
size of the image?

To answer those questions, we take an example of neuroscience field where
containers are used on HPC clusters.

@@

Docker images are organized as a series of layers stacked on top of each
other and identified by unique layer identifiers. If a layer is reused
in different images it will
have the same id across images. Each layer consists of a filesystem diff
with respect to the layer below. Layers are compressed \TG{Are you
referring to file compression similar to gzip?} into a single image and can
be pushed to DockerHub. There are two ways to create a Docker image. The
first is to use an existing Docker image as a base image and then build a
new image from that one, by making changes in the filesystem, which is
saved as a new layer of the Docker image. The second way is to build a new
Docker image from scratch. Docker supports several different storage drivers.
\textit{Overlay2} is the preferred storage driver used by Docker for all currently
supported Linux versions, whereas \textit{aufs} is preferred for Docker 18.06 and
older, when using Ubuntu 14.04 on kernel 3.13 as it do not support \textit{overlay2}.
Other storage drivers that are supported by Docker include \textit{devicemapper, btrfs,
zfs, and vfs}
